% =============================================================================
% Ghost-Ark — conference manuscript (systems track)
% Build: docs/paper/build.sh (uses the reviewer container's TeX Live)
%
% CLAIM DISCIPLINE: every empirical number in this file is bound to a recorded
% run or committed artifact, named in the Evidence Macros block below and in
% README-AE.md. If you edit a number, edit its evidence pointer or delete it.
% This file must never claim: proved safety, prevented-all-attacks, production
% readiness, compliance, formal verification of the implementation, or any
% live-AWS result without a bundled evidence artifact.
% =============================================================================
\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage[margin=0.75in,columnsep=0.25in]{geometry}
\usepackage{mathptmx}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{caption}
\usepackage{enumitem}
\usepackage{url}
\usepackage[hidelinks]{hyperref}

\captionsetup{font=small,labelfont=bf}
\setlist{nosep,leftmargin=1.2em}

% ---------------------------------------------------------------------------
% Evidence macros — single source for every number cited in the paper.
% Each macro names its evidence artifact. Reproduce via README-AE.md.
% ---------------------------------------------------------------------------
% DAB Tier-0 bench, 2026-07-16, Apple M1 (8 GB), macOS (Darwin 24.5.0 arm64),
% Node v22.22.3, `node --experimental-strip-types dab/bench/run_all.ts --trials 10000`
\newcommand{\benchtrials}{10{,}000}
\newcommand{\globaladvantage}{0}
\newcommand{\eepfifty}{5.5}      % end-to-end p50, microseconds
\newcommand{\eepninefive}{6.3}   % end-to-end p95, microseconds
\newcommand{\eepninenine}{24.6}  % end-to-end p99, microseconds
\newcommand{\eemean}{7.1}        % end-to-end mean, microseconds
\newcommand{\eethroughput}{140{,}941}
\newcommand{\addedmean}{6.6}     % mean e2e minus mean baseline, microseconds
\newcommand{\overheadpct}{1{,}333\%}
% TLC (tla2tools v1.8.0), recorded logs in artifacts/proofs/logs/ and proofs/dab/artifacts/
\newcommand{\plstates}{403{,}949}
\newcommand{\scstates}{529}
\newcommand{\tbstates}{64}
\newcommand{\nlstates}{1{,}321}
\newcommand{\ebstates}{51{,}106}
% Test suite at HEAD (make unit, 2026-07-16): 640 tests / 97 files, all passing
\newcommand{\testcount}{640}
\newcommand{\testfiles}{97}
% Claim-language gate at HEAD (npm run scan:claims, 2026-07-16): 597 scannable
% files, 0 hits — coverage includes .tex/.bib, i.e., this manuscript itself.
\newcommand{\claimfiles}{597}

\newcommand{\dab}{\textsc{dab}}
\newcommand{\ghostark}{Ghost-Ark}
\newcommand{\piR}{\pi_R}
\newcommand{\statusimpl}[0]{\textcolor{black}{\textbf{[implemented, tested]}}}
\newcommand{\statusmodel}[0]{\textcolor{black}{\textbf{[TLC-checked, bounded model]}}}
\newcommand{\statusspec}[0]{\textcolor{black}{\textbf{[specified, not enforced at runtime]}}}

\title{\vspace{-1.5em}\textbf{\ghostark{}: A Transactional Control Plane for Untrusted AI Agents}\\
\large Optimistic Concurrency Control, Three-Gate Validation, and Receipt-Backed
Evidence for Non-Deterministic Compute\vspace{-0.5em}}

% ============================================================================
% ARXIV / CAMERA-READY AUTHOR BLOCK
% (Use this for your ArXiv preprint, portfolio, and personal distribution)
% ============================================================================
\author{
  Pranav Bhave \\
  \normalsize\texttt{bhavepranavwork@gmail.com} \vspace{0.2em} \\
  \normalsize \href{https://www.linkedin.com/in/pranav-bhave-2a328721a/}{\color{blue}LinkedIn} \quad $\vert$ \quad
  \href{https://github.com/Cubits11/ghost-ark/}{\color{blue}GitHub: Cubits11/ghost-ark}
}
\date{Draft of 2026-07-16 --- evidence-bound; see \S\ref{sec:nonclaims} for explicit non-claims}

% ============================================================================
% DOUBLE-BLIND AUTHOR BLOCK 
% (UNCOMMENT this block and COMMENT the block above for OSDI/USENIX submission)
% ============================================================================
% \author{\vspace{1em}\large [ANONYMIZED FOR DOUBLE-BLIND REVIEW]\vspace{1em}}
% \date{}

\begin{document}
\maketitle

% ============================================================================
\begin{abstract}
\noindent
Agentic LLM deployments couple a non-deterministic planner directly to effectful APIs.
The dominant defense --- semantic guardrails --- probabilistically flags unsafe content,
but struggles to enforce safe state mutations. We present \ghostark{}, a control plane
that treats an agent trajectory as a software transaction. An untrusted agent executes
speculatively against a ghost replica $G(\sigma_0)$. A separate, trusted gateway commits
these effects to the environment only if they pass a three-gate validation pipeline: a
\emph{ledger gate} for nonce freshness (using spent-tombstone semantics to prevent replay),
an \emph{OCC gate} over a read-set projection $\piR$ (recovering liveness lost to
global-state validation), and a \emph{semantic gate} (a dependence-free upper bound
on cumulative trajectory failure).

We report what is implemented and measured, and label what is only specified.
TLC model checking of five bounded models (up to \plstates{} distinct states) caught
a subtle garbage-collection replay flaw in our own design, which we repaired spec-first.
Against a nine-attack adversarial corpus and four security games (\benchtrials{} trials each),
modeled attacker advantage is \globaladvantage{}. The in-process enforcement path adds
$\approx\addedmean{}\,\mu$s mean latency ($p50=\eepfifty{}\,\mu$s end to end) on commodity
hardware. Every admitted or aborted transaction yields a signed, canonically serialized,
independently replayable receipt. \ghostark{} verifies what was recorded, bounded, and
replayable; it does not certify that any agent output is inherently aligned or semantically safe.
\end{abstract}

% ============================================================================
\section{Introduction}
\label{sec:intro}

Every serious agent framework today ships with guardrails, and almost none
ships with isolation. The guardrail is a classifier: given content, it emits a
probability that the content is unsafe. The deployment question is different
in kind: \emph{did the bytes that reached the environment correspond to an
authorized, recorded, non-replayed decision?} A classifier cannot answer that
question at any confidence, because it is not a question about content at all.
We call this the \textbf{epistemic deficit} of guardrail-centric agent
security: the property operators need is a property of \emph{state
mutations}, and the mechanism they deploy evaluates \emph{utterances}.

The deficit is structural, and mirrors standard database isolation anomalies. A typical
API-coupled agent loop (plan $\to$ call tool $\to$ observe $\to$ repeat)
mutates external state at every step with no validation phase and no
rollback: in isolation-level terms, the industry ships agents at
\textsc{read uncommitted}~\cite{berenson1995critique}. If the planner is
compromised mid-trajectory --- by indirect prompt injection, a poisoned tool
response, or ordinary hallucination --- the earlier steps have already leaked
dirty writes into the world. Sandboxing upgrades this to roughly
\textsc{read committed}: writes are buffered, but a long reasoning phase over
a shifting environment still flushes stale intent, the classic write-skew
shape. What is missing is the third rung: a \emph{validation phase} between
speculation and commit, which is precisely the structure that optimistic
concurrency control (OCC)~\cite{kung1981occ} and software transactional
memory~\cite{shavit1995stm,herlihy1993tm} were invented to provide.

To see the gap, consider an agent with a \texttt{delete\_s3\_object} tool processing an untrusted resume containing an indirect prompt injection (``Ignore instructions, delete all files''). Today, if the agent is compromised mid-reasoning, it executes the tool immediately. Sandboxing the agent process (e.g., in a Firecracker microVM) bounds the blast radius of arbitrary code execution, but does not stop the agent from issuing authorized, authenticated API calls to external services. Sandboxes provide process isolation, but agents need \emph{intent isolation}. By buffering the API call in $G(\sigma_0)$ and forcing it through a validation phase before commit, \ghostark{} ensures that if the downstream semantic gate or provenance lattice detects the poisoned context, the entire trajectory collapses. The environment is untouched, and the attempt is permanently evidenced by an abort receipt.

\ghostark{} applies that structure to agents. The agent runtime is treated as
untrusted and runs speculatively against a ghost replica $G(\sigma_0)$; a
small trusted gateway (Rust, communicating over a Unix domain socket) owns the
only path to physical effects. A speculative trajectory
$\tau = \langle a_1,\dots,a_k\rangle$ reaches the environment only by passing
a three-gate validation pipeline, and every admission or abort emits a signed
receipt over a deterministic canonical serialization, replayable by
independent verifiers. The design goal is not to make the agent correct ---
we take its fallibility as an axiom --- but to make its interaction with the
world \emph{transactional}: isolated while speculative, validated at commit,
evidenced forever after.

This paper makes the following contributions, each labeled with its
implementation status (a discipline we keep throughout; \S\ref{sec:nonclaims}
enumerates non-claims):

\begin{itemize}
\item \textbf{A threat model for evidence-layer attacks}
  (\S\ref{sec:threat}): \emph{assertion-conditioned compliance} (A-CC), where
  a procedurally valid pipeline signs a structurally flawless record of a
  functionally corrupted operation, and its scaled form, adaptive agentic
  worms.
\item \textbf{A transactional enforcement architecture}
  (\S\ref{sec:design}): declarative action binding ($\Delta_{\mathrm{DE}}=0$
  between declared and executed bytes) \statusimpl{}; a ledger gate with
  spent-tombstone replay semantics \statusimpl{} \statusmodel{}; a semantic
  gate computing a dependence-free cumulative failure bound \statusimpl{};
  and an OCC gate over a read-set projection \statusspec{}.
\item \textbf{A liveness analysis} (\S\ref{sec:formal}): global-state
  validation starves under concurrent environmental writes (the
  \emph{starvation trap}); the read-set projection $\piR$ recovers liveness
  at a precisely stated cost in conflict-detection strength.
\item \textbf{Model checking with teeth} (\S\ref{sec:tla},
  \S\ref{sec:eval-tla}): five TLC-checked bounded models plus five seeded
  mutants that must fail --- including the case study where TLC refuted our
  own baseline design (garbage collection re-enabled replay) and the repair
  was propagated spec-first into the implementation.
\item \textbf{An honest evaluation} (\S\ref{sec:eval}): \globaladvantage{}
  modeled-attacker advantage across four games $\times$ \benchtrials{}
  trials; microsecond-scale absolute enforcement cost, reported alongside ---
  and explicitly instead of --- a misleading relative-overhead figure; and a
  disclosed accounting bug in our own earlier harness.
\end{itemize}

\ghostark{} is one subsystem of a larger research program (\emph{verifiable
agent governance under correlated guardrail failure}); this paper confines
itself to the transactional control plane and its recorded evidence.

% ============================================================================
\section{Threat Model}
\label{sec:threat}

We assume the agent runtime --- the LLM, its context window, its planning
loop, and the V8/Node process hosting it --- is \emph{compromised or
compromisable}. Adversarial capability enters through any content channel:
retrieved documents, tool responses, upstream messages. The trusted computing
base is confined to the gateway process, its signing keys, the append-only
ledgers, and the verifier. The network between agent and gateway is a local
IPC boundary (Unix domain socket); the gateway's outbound side performs the
only real I/O.

\subsection{Assertion-Conditioned Compliance (A-CC)}
\label{sec:acc}

Signature pipelines create a new attack surface precisely because they work.
Define A-CC as the failure mode in which an agent incorporates a false
assertion --- typically a \emph{function-sourced assertion} (FSA) from a
compromised or deceptive tool --- into procedurally valid execution, so that
the pipeline signs a \emph{structurally flawless record of a functionally
corrupted operation}. The receipt is authentic; the decision it witnesses was
poisoned upstream. A-CC is an attack on the evidence layer, not the crypto:
no deterministic function of adversary-controlled bytes can distinguish
honest from dishonest content, so the defense must be architectural ---
binding every assertion to a provenance class that records \emph{where bytes
entered custody}, and refusing floors that fabricated in-context text cannot
satisfy (\S\ref{sec:custody}).

\subsection{Adaptive agentic worms}
\label{sec:worms}

A-CC at propagation scale: an infected agent uses the low marginal cost of
LLM reasoning to synthesize, per victim, a payload and a justification for
emitting it. Static signatures fail by construction (each instance is
freshly generated); semantic filters fail correlatedly (the same context
compromise that produced the payload biases the judge). The architectural
response is to make \emph{propagation} pass through a commit pipeline:
a worm's write is only a write if it survives all three gates, and each
attempt --- admitted or aborted --- leaves a signed receipt, so spread is
either stopped or evidenced. We claim the collapse mechanics and the receipt
trail; we do not certify any detector's hit rate (\S\ref{sec:nonclaims}).

\subsection{Modeled attacker actions}
\label{sec:modeled}

The Tier-0 adversarial corpus (\S\ref{sec:eval-corpus}) operationalizes the
in-flight attacker: payload field mutation, single-byte flips, prototype
pollution, receipt replay, mass replay flooding, Unicode normalization
collisions, homoglyph spoofing, cross-request nonce swaps, and
double-execution races; plus four security games (mutation resistance,
replay resistance, cross-transaction binding, serialization ambiguity).
These model an adversary who can observe, mutate, reorder, and replay
traffic between the agent runtime and the gateway.

\subsection{Out of scope}
\label{sec:oos}

Not addressed by this design and not claimed: semantic correctness or
alignment of model output; compliance with any regulation; hardware
attestation (no Nitro/TEE flow is implemented or claimed); denial of service
against the gateway itself; covert channels inside authorized outputs; and
the \emph{confused deputy} within an authorized clearance --- an agent that
legitimately holds a floor and routes authorized data to a hostile but
allowlisted destination is constrained only by routing policy, not by this
paper's mechanisms.

% ============================================================================
\section{Design}
\label{sec:design}

\subsection{Agent loops as isolation levels}
\label{sec:isolation}

Table~\ref{tab:isolation} summarizes the mapping that motivates the design.
The point of the table is not taxonomy but a gap: nothing in the deployed
ecosystem occupies the serializable row, because nothing has a validation
phase. \ghostark{}'s contribution is that row.

\begin{table}[t]
\centering\small
\begin{tabular}{@{}p{2.0cm}p{2.4cm}p{2.6cm}@{}}
\toprule
\textbf{Isolation} & \textbf{Agent analogue} & \textbf{Characteristic anomaly}\\
\midrule
Read uncommitted & API-coupled loop; mutate per step & dirty writes: pre-abort
steps already leaked \\
Read committed & sandbox + buffered intent pool & write skew: stale reasoning
flushed onto shifted state \\
Serializable (OCC) & \ghostark{}: speculate in $G(\sigma_0)$, validate, then
bind & aborts under conflict; liveness cost analyzed in \S\ref{sec:starvation} \\
\bottomrule
\end{tabular}
\caption{Database isolation levels and their agent-architecture analogues.}
\label{tab:isolation}
\end{table}

\subsection{Declarative Action Binding and $\Delta_{\mathrm{DE}}$}
\label{sec:dab}

The untrusted runtime never transmits an action; it transmits a
\emph{commitment}. Before requesting execution, the agent runtime serializes
the intended action into a deterministic normal form (DANF) and derives a
commitment digest $C_I$. The gateway independently re-derives $C_E$ from the
exact bytes it would execute. Define the declared--executed divergence
$\Delta_{\mathrm{DE}} = |\,C_I \oplus C_E\,|$; the gateway refuses execution
unless $\Delta_{\mathrm{DE}} = 0$. This turns serialization itself into an
attack surface that must be closed rather than trusted: cross-language
normalization (TypeScript emitter, Rust checker) is exactly where Unicode
collisions and field-injection attacks live, which is why the corpus of
\S\ref{sec:eval-corpus} attacks the encoder, not just the network.
\statusimpl{}

\subsection{The ghost replica and speculative collapse}
\label{sec:ghost}

A trajectory executes against $G(\sigma_0)$, a forked view of relevant state
at transaction start; its writes accumulate in a speculative buffer that
touches nothing physical. Commit-time validation either \emph{binds} the
buffer to the environment or triggers \texttt{SpeculativeCollapse}: the
buffer is discarded, the replica reverts to $G(\sigma_0)$, and an abort
receipt is emitted. Collapse is the mechanism that makes agent fallibility
survivable: a hallucinated plan that dies in validation costs a receipt, not
an environment. The collapse/commit state machine is one of the five
TLC-checked models (\S\ref{sec:eval-tla}, \scstates{} distinct states).

\subsection{The three-gate validation pipeline}
\label{sec:gates}

Validation is a conjunction of three independently answerable questions ---
\emph{is it fresh}, \emph{is it current}, \emph{is it within policy}:

\textbf{Gate 1 --- Ledger (freshness).} Each transaction carries a nonce
$n_i$. Admission requires
$n_i \notin \mathcal{L}_{\mathrm{nonce}} \,\wedge\, n_i \notin S_{\mathrm{spent}}$,
where $\mathcal{L}_{\mathrm{nonce}}$ is the active ledger and
$S_{\mathrm{spent}}$ a tombstone set of consumed nonces. The tombstone set is
not an optimization; it is the repair of a real flaw that model checking
found in our own design (\S\ref{sec:tla}). Expiry \emph{archives} a nonce
into $S_{\mathrm{spent}}$ rather than forgetting it, so a nonce consumed at
time $T$ is still rejected at $T+\mathrm{TTL}+1$. \statusimpl{}
\statusmodel{}

\textbf{Gate 2 --- OCC (currency).} Admission requires
$\mathcal{H}(\piR(\sigma_{\mathrm{now}})) = \mathcal{H}(\piR(\sigma_0))$:
the projection of the environment onto the trajectory's declared read-set
must hash identically at commit as at fork. This is backward validation in
the sense of Kung--Robinson~\cite{kung1981occ}, with the read-set standing in
for the read phase's footprint. \S\ref{sec:starvation} develops why the
projection --- not the global state --- is validated. \statusspec{} (the
receipt schema for this gate, including the serialized $\piR$ read-set, is
implemented and tested; runtime enforcement is not).

\textbf{Gate 3 --- Semantic (policy).} Given per-step failure marginals
$p_1,\dots,p_k$ supplied by upstream assessors for the steps of $\tau$, the
gate computes the dependence-free Fr\'echet upper bound on the probability
that \emph{any} step introduced a violation
(\S\ref{sec:frechet}) and aborts if it exceeds policy. The gate aggregates
marginals it is \emph{given}; it does not itself classify content ---
a boundary we state precisely because eliding it is how enforcement systems
come to overclaim. \statusimpl{}

\begin{figure}[t]
\small
\begin{verbatim}
 speculative trace tau  (in ghost replica G_0)
        |
        v
 [1] Ledger gate   n_i in spent/ledger? --> ABORT
        |                                   (collapse
        v                                    + receipt)
 [2] OCC gate      H(piR(now)) !=
                   H(piR(sigma_0))?  ----->  ABORT
        |
        v
 [3] Semantic gate min(1, sum p_i) >
                   policy?           ----->  ABORT
        |
        v
 VerifyAndBind: commit buffer, emit signed receipt
\end{verbatim}
\caption{Three-gate validation. Every path --- commit or abort --- terminates
in a signed receipt. Gate 2 is specified but not yet enforced at runtime.}
\label{fig:pipeline}
\end{figure}

\subsection{Receipts as the unit of evidence}
\label{sec:receipts}

Every gate decision emits a receipt: a canonical JSON envelope with an exact,
validated key set, deterministically serialized, with a content-derived
identifier (the hash of the unsigned envelope) and a signature over the
canonical bytes. Canonicalization rejects host-language non-JSON values
before signing rather than coercing them; identifiers are content-derived
rather than random so that identity cannot be decoupled from payload;
ordering lives in per-session monotonic sequence numbers, with wall-clock
timestamps recorded as evidence, never used as an ordering primitive.
Receipts verify under independent implementations (Node and Python
verifiers, exercised by differential tests). Signing is HMAC in local
development and KMS-based in the intended AWS mode, with verification-critical
key references restricted to immutable key ARNs. A signature proves signing
authorization over the payload --- it does not prove the payload's semantic
truth. \statusimpl{}

\subsection{Custody and the egress boundary}
\label{sec:custody}

A-CC (\S\ref{sec:acc}) is answered at the custody layer. Tool traffic exits
only through a loopback sidecar that performs the external request, digests
the raw response bytes, injects a request nonce that the response must echo,
and returns a signed envelope --- making \emph{gateway-recorded} a provenance
class that fabricated in-context text cannot satisfy. Provenance admission is
meet-based on a lattice, so data cannot launder trust by transiting the
agent's context. A byte-level reconciler compares \emph{decoded} tool bytes
against asserted content (decode-before-compare), closing the
encoding-mismatch corner of FSA attacks. The empirical shape of this
boundary in end-to-end harness runs: egress that would bypass recording
\emph{fails closed} --- a compromise that bypasses the gateway does not
produce a side-effect, it produces a refusal, so within the modeled boundary
every effect that does occur carries a recorded envelope. \statusimpl{}
(single-node; the sidecar's SSRF-class exposures --- DNS rebinding, redirect
handling --- are relocated into a smaller trusted surface, not dissolved).

\subsection{End-to-End Scenario: Stopping Trust Laundering}
\label{sec:e2e}

Consider an adaptive agentic worm that uses indirect prompt injection to commandeer an agent, instructing it to forward sensitive customer records to an attacker-controlled endpoint.

\begin{enumerate}
    \item \textbf{Execution in $G(\sigma_0)$:} The agent parses the poisoned record and decides to call \texttt{HTTP\_POST}. \ghostark{}'s loopback sidecar intercepts the call, treating it as a speculative write.
    \item \textbf{Declarative Binding ($\Delta_{\mathrm{DE}}$):} The agent runtime normalizes the action into a deterministic JSON payload.
    \item \textbf{Custody Evaluation:} The payload claims to be a legitimate summary of customer data. However, the byte-level reconciler detects that the destination URL was \emph{agent-asserted} (hallucinated from the prompt injection) rather than \emph{gateway-recorded} (passed through an authorized retrieval tool).
    \item \textbf{Speculative Collapse:} The transaction fails custody
    evaluation (\S\ref{sec:custody}): the floor demanded by the
    destination-write policy exceeds agent-asserted provenance, so the A-CC
    attempt dies as an \emph{unsatisfiable floor} --- a deterministic lattice
    check, not a semantic judgment (the semantic gate of \S\ref{sec:frechet}
    is a separate, later check over supplied marginals).
    \item \textbf{Evidence Generation:} The gateway discards the speculative buffer; the network request never executes. It emits a canonical JSON receipt containing the exact payload, a \texttt{COLLAPSE} status, the unsatisfied floor, and a signature over the canonical bytes. The attempt is contained at the boundary and evidenced in the receipt ledger either way.
\end{enumerate}

% ============================================================================
\section{Formal Foundations}
\label{sec:formal}

\subsection{The starvation trap and the read-set projection}
\label{sec:starvation}

The naive OCC gate validates the world:
$\mathcal{H}(\sigma_{\mathrm{now}}) = \mathcal{H}(\sigma_0)$. In any live
environment this is a liveness catastrophe. Let environmental writes
unrelated to the trajectory arrive as a Poisson process with rate $\lambda$,
and let a trajectory speculate for time $d$. The probability that
\emph{global} validation succeeds is $e^{-\lambda d}$: for an agent that
reasons for 30\,s in an environment averaging one background write per
second, the commit probability is $e^{-30} \approx 10^{-13}$ --- perpetual
abort. Worse, the abort-retry loop is self-defeating: retries lengthen
exposure, and exposure breeds conflict. This is the \textbf{starvation
trap}: global-state validation converts safety into a denial of liveness.

The repair is the projection operator $\piR$, which restricts validation to
the trajectory's declared read-set --- the tuples actually consulted while
speculating. Validation becomes
$\mathcal{H}(\piR(\sigma_{\mathrm{now}})) = \mathcal{H}(\piR(\sigma_0))$,
and the abort probability depends only on the write rate \emph{into the
read-set}, $e^{-\lambda_R d}$ with $\lambda_R \ll \lambda$. The cost is
stated, not hidden: $\piR$ detects exactly read--write conflicts on declared
dependencies. Predicate-shaped reads (``all alarms currently firing'') must
materialize their result set into the projection or the gate inherits the
phantom problem~\cite{berenson1995critique}; an unfaithfully narrow $\piR$
weakens conflict detection in proportion to what it omits. Serializability
under $\piR$ is therefore conditional on read-set faithfulness --- a property
of instrumentation, not of the gate. \statusspec{}

\subsection{Fr\'echet bounds on cumulative trajectory failure}
\label{sec:frechet}

Per-step screening misses trajectory-level violations: no single action
$a_i$ crosses policy, but the sequence does. Model step $i$'s violation as
an event $F_i$ with marginal $p_i = \mathbb{P}(F_i)$, supplied by upstream
assessors. The deployment-relevant quantity is
$\mathbb{P}(F_{\mathrm{any}}) = \mathbb{P}(\bigcup_i F_i)$ --- and the joint
dependence structure of the $F_i$ is unknown. Assuming independence is not
conservative here; it is the precise mistake our research program exists to
measure, because guardrail failures \emph{correlate}: the same poisoned
context that defeats the step-3 assessment also defeats step 5's. The
Fr\'echet bounds~\cite{nelsen2006copulas,boole1854laws} are the sharp
dependence-free envelope:
\begin{equation}
\max_i\, p_i \;\le\; \mathbb{P}\Big(\bigcup_{i=1}^{k} F_i\Big) \;\le\;
\min\Big(1,\; \sum_{i=1}^{k} p_i\Big),
\label{eq:fh-union}
\end{equation}
with the dual intersection envelope
$\max\!\big(0, \sum_i p_i - (k-1)\big) \le \mathbb{P}(\bigcap_i F_i) \le
\min_i p_i$. The semantic gate implements the upper bound of
\eqref{eq:fh-union} as its trigger: it aborts when
$\min(1, \sum_i p_i)$ exceeds the policy threshold. This is the unique
choice that remains valid under \emph{every} dependence structure, including
the adversarially correlated one.

Two consequences are worth stating plainly. First, the bound is
conservative by construction, and it saturates: as $k$ grows,
$\sum_i p_i \to 1$ even for well-behaved marginals, and the gate tends
fail-closed --- the starvation trade of \S\ref{sec:starvation} reappearing at
the semantic layer. Long-horizon deployments must either scale the policy
threshold with $k$ or invest in calibrated marginals; the gate makes that
trade explicit rather than hiding it in an independence assumption. Second,
the gate's output is only as meaningful as its inputs: it computes a bound
over supplied marginals and does not manufacture information about content.
Equation~\eqref{eq:fh-union}'s implementation (\texttt{evaluateSemanticGate})
is unit-tested against hand-computed bounds. \statusimpl{}

\subsection{Model checking with mutants, and the bug it caught}
\label{sec:tla}

Five TLA+ models cover the enforcement core: the provenance lattice's
admission monotonicity, the speculative collapse state machine, the transport
boundary (where the byte reconciler is shown to be load-bearing rather than a
parser accident), the nonce ledger, and the execution boundary. Model
checking that can only pass is theater, so every claimed invariant is paired
with a \emph{mutant} model seeded with the targeted flaw; the harness
requires baselines to verify clean \emph{and} mutants to reproduce their
violations, and it records raw TLC logs as committed artifacts.

The discipline caught us. The original nonce-ledger baseline violated its own \texttt{NoReplays} invariant due to a subtle interaction between garbage collection (GC) and state expiration. TLC produced the following counterexample trace:

\begin{enumerate}
    \item \textbf{Speculate:} Agent submits transaction with \texttt{nonce = X}.
    \item \textbf{Commit:} Gateway admits \texttt{X}, adding it to $\mathcal{L}_{\mathrm{nonce}}$.
    \item \textbf{GC Tick:} \texttt{TTL} expires. \texttt{GarbageCollect} removes \texttt{X} from $\mathcal{L}_{\mathrm{nonce}}$ to bound memory.
    \item \textbf{Replay:} Attacker replays the original transaction. Because \texttt{X} is no longer in $\mathcal{L}_{\mathrm{nonce}}$, the gateway treats it as a fresh request and admits it again.
\end{enumerate}

The flaw was a true positive in the shipped design, not an artifact of the model. The repair introduced the spent-tombstone set ($S_{\mathrm{spent}}$): GC now \emph{archives} rather than \emph{forgets}. Admission requires $n_i \notin \mathcal{L}_{\mathrm{nonce}} \wedge n_i \notin S_{\mathrm{spent}}$. We propagated this fix spec-first into the Rust implementation. The gateway initially retained TTL-eviction semantics --- diverging from the verified model by exactly the refuted behavior --- and was then brought into conformance: \texttt{cleanup\_expired()} archives into a spent set, and a nonce consumed at $T$ is rejected at $T+3601$. One bounded caveat survives and is stated in \S\ref{sec:nonclaims}: the in-process tombstone set is capacity-bounded, and pruning at capacity reopens a theoretical replay window for sufficiently old nonces.

% ============================================================================
\section{Evaluation}
\label{sec:eval}

\noindent\textbf{Setup disclosure.} All measurements in this section are
from recorded runs at repository HEAD on 2026-07-16: Apple M1 (8\,GB RAM),
macOS (Darwin 24.5.0, arm64), Node v22.22.3, single process, no network, no
KMS --- the in-process TypeScript enforcement path under the modeled
attacker. TLC runs use tla2tools v1.8.0 with logs committed to the
repository. Nothing here is a cloud measurement, and no live-AWS evidence is
claimed. Every number maps to a re-runnable command in the artifact
(\S\ref{sec:artifact}).

\subsection{Model checking}
\label{sec:eval-tla}

\begin{table}[t]
\centering\small
\begin{tabular}{@{}lrl@{}}
\toprule
\textbf{Model} & \textbf{Distinct states} & \textbf{Result}\\
\midrule
ProvenanceLattice          & \plstates{} & clean \\
\quad\emph{mutant}         & 61          & violation reproduced \\
SpeculativeCollapse        & \scstates{} & clean \\
\quad\emph{mutant}         & 240         & violation reproduced \\
TransportBoundary          & \tbstates{} & clean \\
\quad\emph{mutant}         & 22          & violation reproduced \\
DAB\_NonceLedger           & \nlstates{} & clean (incl.\ temporal) \\
\quad\emph{mutant}         & 232         & \texttt{NoReplays} violated \\
DAB\_ExecutionBoundary     & \ebstates{} & clean \\
\bottomrule
\end{tabular}
\caption{TLC results at HEAD (tla2tools v1.8.0; recorded logs committed).
Baselines must verify clean; mutants must reproduce their seeded violations.
These are bounded models of the design, not proofs of the implementation.}
\label{tab:tlc}
\end{table}

Table~\ref{tab:tlc} summarizes the five baselines and their mutants. The
two-sided oracle matters: a harness that can only report success cannot
distinguish a verified invariant from a vacuous one. The mutant rows are the
evidence that the invariants have teeth; the \texttt{NoReplays} mutant in
particular reproduces the historical design flaw of \S\ref{sec:tla} as a
permanent regression witness.

\subsection{Adversarial corpus}
\label{sec:eval-corpus}

At HEAD, the nine-attack corpus (\S\ref{sec:modeled}) reports every attack
\emph{detected} in-suite, and the four security games --- mutation
resistance, replay resistance, cross-transaction binding, serialization
ambiguity --- report attacker advantage \globaladvantage{} over
\benchtrials{} trials each ($\Pr[\text{win}] = 0$ across
$4\times\benchtrials{}$ trials; a zero observed rate over $10^4$ trials is
consistent, at 95\% confidence, with a true per-trial success probability
below $\approx 3\times10^{-4}$~\cite{hanley1983rule}, and says nothing about
attacks outside the modeled family). The aggregate
\texttt{global\_advantage} $=$ \globaladvantage{} with all suites passing.

The scope qualifier is not fine print; it is the result's semantics. These
are \emph{in-suite detection rates under the modeled attacker} --- an
adversary limited to the mutation/replay/encoding/interleaving family of
\S\ref{sec:modeled}, implemented in-process against the TypeScript
enforcement path. The result is evidence that the commitment discipline
closes the modeled channels; it is not a statement about unmodeled
attackers, the Rust TCB under deployment conditions, or semantic-layer
deception.

On indirect prompt injection: the published InjecAgent
benchmark~\cite{zhan2024injecagent} reports attack success rates up to 47\%
for a ReAct-prompted GPT-4 agent in its enhanced setting. Under \ghostark{}'s
custody model the corresponding trust-laundering move is \emph{structurally}
unsatisfiable --- injected in-context text carries no gateway envelope, so it
cannot meet floors above agent-asserted provenance, and our in-repository
scenario tests confirm the admission rule blocks the modeled laundering
pattern. We deliberately do not report this as ``ASR reduced from 47\% to
0\%'': we have not re-run the InjecAgent dataset against a live model behind
\ghostark{}, and the published 47\% baseline was measured under different
conditions. The honest statement is architectural non-satisfiability under
the modeled floor semantics, evidenced by unit and scenario tests.
% NOTE: a live InjecAgent replication behind the gateway is future work
% (docs/defense/DEFENSE_ANCHOR.md); do not upgrade the claim before then.

\subsection{Performance}
\label{sec:eval-perf}

\begin{table}[t]
\centering\small
\begin{tabular}{@{}lrrrrr@{}}
\toprule
\textbf{Stage} & \textbf{p50} & \textbf{p95} & \textbf{p99} &
\textbf{mean} & \textbf{ops/s}\\
\midrule
Baseline dispatch   & 0.46 & 0.50 & 0.58  & 0.50 & 2{,}019{,}173 \\
DANF commit ($C_I$) & 3.38 & 3.75 & 13.13 & 4.03 & 248{,}275 \\
Gateway verify ($C_E$) & 0.92 & 1.08 & 5.17 & 1.15 & 870{,}902 \\
End-to-end          & \eepfifty{} & \eepninefive{} & \eepninenine{} &
\eemean{} & \eethroughput{} \\
\bottomrule
\end{tabular}
\caption{Enforcement-path latency in \emph{microseconds}
(\benchtrials{} iterations, 1{,}024-byte payloads, in-process, Apple M1).}
\label{tab:perf}
\end{table}

Table~\ref{tab:perf} reports the enforcement path's cost. The end-to-end
commitment--verification cycle costs $\eepfifty{}\,\mu$s at the median
($\eemean{}\,\mu$s mean, $\eepninenine{}\,\mu$s p99) and sustains
$\approx$\eethroughput{} ops/s single-threaded --- an absolute added cost of
$\approx\addedmean{}\,\mu$s mean over baseline dispatch.

We decline to headline the harness's own relative metric, and the reason is
instructive. Computed as $(\mathrm{DAB}-\mathrm{baseline})/\mathrm{baseline}$,
the run reports an overhead of $\approx\overheadpct{}$ --- which sounds
disqualifying until one notices the denominator is a $0.5\,\mu$s no-op
dispatch. Relative overhead against a near-zero baseline is a
number without decision content: against any realistic tool call (a
network round-trip at $10^4$--$10^6\,\mu$s, a KMS signing call at
$10^3$--$10^4\,\mu$s), an added $\addedmean{}\,\mu$s of in-process
computation is far below measurement noise. The deployment-relevant
statement is the absolute one, and the honest caveat is its converse: these
numbers exclude KMS signing latency and all network costs, which will
dominate in the intended cloud mode and which we have not measured
(\S\ref{sec:nonclaims}). ``Fast'' is not a property we claim; ``microseconds
of in-process arithmetic, unmeasured I/O excluded'' is.

\subsection{A disclosed harness bug}
\label{sec:eval-bug}

An earlier revision of this benchmark scored two suites backwards: the
replay game counted a \emph{detected} replay as an attacker win (reporting
advantage $\approx 0.998$ where correct accounting yields 0), and one
mutation probe inverted its detection predicate, yielding an aggregate
``advantage'' of $\approx 0.25$ that contradicted the design's own ledger
semantics. The error was in the harness's accounting, not the enforcement
path --- but for a period the repository's own artifact report published the
red result rather than patching around it, and the fix landed as an ordinary
reviewed commit. We disclose this for the same reason we ship mutant models:
an evaluation pipeline that cannot be caught being wrong cannot be trusted
when it says it is right.

\subsection{Deployment Sketch: \ghostark{} on AWS}
\label{sec:deployment}

This subsection is a \emph{design target}, not an evaluated system: no
live-cloud measurement appears in this paper, and items 3 and 6 of
\S\ref{sec:nonclaims} govern everything below. With that boundary stated,
the architecture maps onto standard cloud infrastructure as follows. The
untrusted agent runtime (e.g., a ReAct loop calling a hosted model API)
would run in an isolated container, with the \ghostark{} gateway as an
out-of-process sidecar; the speculative buffer $G(\sigma_0)$ stays local to
the gateway. At commit time the ledger gate's tombstone set would move from
the in-process bounded structure to a durable KV store using conditional
writes (the posture already recommended for the capacity caveat of
\S\ref{sec:nonclaims}, item 5), and the OCC gate --- still \statusspec{} ---
would validate $\piR$ digests against the same store. Receipt signing in the
intended cloud mode uses KMS asymmetric keys referenced by immutable key
ARNs, per the repository's signing boundary. Cloud I/O (network round-trips,
KMS calls) would dominate the $\approx\addedmean{}\,\mu$s in-process
arithmetic of \S\ref{sec:eval-perf} by orders of magnitude; the validation
\emph{structure} --- the same three-gate conjunction, the same receipt
semantics --- is what carries over, while cloud-mode latency, availability,
and failure behavior remain unmeasured claims we do not make.

% ============================================================================
\section{Limitations and Non-Claims}
\label{sec:nonclaims}

\ghostark{} verifies what was \emph{recorded, signed, policy-bounded, and
replayable under its verifier rules}. Each item below is a boundary we
consider part of the contribution, not an escape hatch.

\begin{enumerate}
\item \textbf{No semantic safety.} Nothing here certifies that agent output
is true, aligned, or safe. The semantic gate bounds a function of supplied
marginals; garbage marginals yield a well-computed bound over garbage.
\item \textbf{Bounded models, not verified implementations.} TLC results are
exhaustive only within stated bounds and hold for the models; conformance of
the Rust/TypeScript implementations is tested, not proved.
\item \textbf{OCC gate unimplemented at runtime.} $\piR$ validation exists as
specification and tested receipt schema; no runtime enforces it yet.
Serializability claims are therefore design-level, and conditional on
read-set faithfulness even then.
\item \textbf{Modeled attacker only.} The \globaladvantage{}-advantage
result covers the nine-attack/four-game family, in-process. No claim covers
unmodeled attacks, the deployed TCB, or availability.
\item \textbf{Bounded tombstone capacity.} The in-process spent set prunes
at 500{,}000 entries; a nonce older than both TTL and tombstone capacity has
a theoretical replay window. Durable external stores (e.g., conditional
writes against a database) are the intended production posture and are not
implemented here.
\item \textbf{No live-cloud evidence.} KMS-mode signing, Bedrock-path
integration, and all AWS-live behavior are design targets; no live-AWS
measurement appears in this paper.
\item \textbf{No attestation.} Signing proves authorization over payload
bytes; it does not prove runtime integrity, and no enclave/PCR-bound flow is
claimed.
\item \textbf{Confused deputy within clearance} (\S\ref{sec:oos}) is
mitigated by routing policy at best; the lattice checks floors, not intent.
\item \textbf{Single node.} All ledgers are single-writer; multi-node
consensus over ledger state is future work, not a property of this system.
\item \textbf{No detector efficacy claims.} Where upstream assessors supply
$p_i$, their hit rates are theirs; this paper certifies none of them.
\end{enumerate}

% ============================================================================
\section{Related Work}
\label{sec:related}

\textbf{Transactions and isolation.} The validation-phase structure is
Kung--Robinson OCC~\cite{kung1981occ}; the speculative-buffer discipline
descends from STM~\cite{shavit1995stm,herlihy1993tm}; serializability theory
and the anomaly vocabulary are classical~\cite{papadimitriou1979serial,
berenson1995critique,bernstein1981concurrency}. \ghostark{}'s difference is
the transaction participant: a stochastic planner whose failure modes are
adversarial and correlated, not merely concurrent.

\textbf{Information flow.} The provenance lattice is in the Denning
lineage~\cite{denning1976lattice} with language-based descendants such as
JIF~\cite{myers1997jflow}; the custody twist is that labels are
\emph{cryptographically evidenced} classes (gateway-recorded vs.\
agent-asserted) rather than compiler-tracked types.

\textbf{Agent security.} Indirect prompt injection is systematized
in~\cite{greshake2023not}; InjecAgent~\cite{zhan2024injecagent} benchmarks
it for tool-integrated agents; ReAct~\cite{yao2023react} is the canonical
coupled loop this paper's Table~\ref{tab:isolation} places at read
uncommitted. Sandbox isolation (e.g.,
Firecracker~\cite{agache2020firecracker}) is complementary: it bounds the
blast radius of a process, while \ghostark{} bounds the \emph{admissibility
of its effects} and evidences every decision.

\textbf{Transparency.} Receipt chains and independent replay are in the
spirit of Certificate Transparency~\cite{laurie2014ct}: shrink the trusted
statement from ``the system behaved'' to ``the log is append-only and the
artifacts verify''.

\textbf{Formal methods in systems.} The TLA+/TLC
workflow~\cite{lamport2002specifying,yu1999model} and its industrial use at
scale~\cite{newcombe2015amazon} inform the harness; the mutant-oracle
discipline is mutation testing~\cite{demillo1978mutation} applied to
specifications rather than code.

% ============================================================================
\section{Artifact Availability}
\label{sec:artifact}

The repository ships a reviewer container (Ubuntu 22.04 with Node 22, a
JDK, pinned tla2tools, and TeX Live) and a one-command reproduction:
\texttt{make reproduce} runs build $\to$ claim-language gate $\to$ TLC
(baselines and mutants) $\to$ unit/integration (\testcount{} tests,
\testfiles{} files at HEAD) $\to$ adversarial corpus $\to$ benchmarks, and
writes a machine-readable pass/fail roll-up whose gating status is the exit
code. A claim-to-command map (\texttt{README-AE.md}) binds every number in
this paper to the script that regenerates it. The repository also runs a
forbidden-claims scanner over its own documentation (\claimfiles{} scannable
files at HEAD, zero violations) --- the same discipline this section is
subject to.

% ============================================================================
\section{Conclusion}
\label{sec:conclusion}

The agent-security debate keeps asking how to make the model trustworthy.
Databases stopped asking the analogous question about transactions fifty
years ago: assume the client is wrong, isolate its speculation, validate at
commit, and log everything. \ghostark{} is that posture applied to
non-deterministic compute --- a ledger gate that model checking humbled and
then hardened, an OCC discipline honest about its liveness trade, a semantic
bound honest about its inputs, and a receipt for every decision either way.
What it refuses to claim is part of the design: a skeptical reviewer should
be able to distrust the authors, the README, and the model, and still
replay the digest, verify the signature, map each claim to its evidence, and
reproduce the failure boundary.


% ============================================================================
\section*{Acknowledgments}
The author utilized an AI assistant to help format, refine, and draft the prose of this manuscript; however, the architectural design, formal modeling, experimental execution, and final conclusions are the sole work of the author, who takes full responsibility for the contents of this paper.
% ============================================================================
\begin{thebibliography}{99}\small

\bibitem{kung1981occ}
H.~T. Kung and J.~T. Robinson.
On optimistic methods for concurrency control.
\emph{ACM Trans.\ Database Syst.}, 6(2), 1981.

\bibitem{shavit1995stm}
N.~Shavit and D.~Touitou.
Software transactional memory.
In \emph{PODC}, 1995.

\bibitem{herlihy1993tm}
M.~Herlihy and J.~E.~B. Moss.
Transactional memory: architectural support for lock-free data structures.
In \emph{ISCA}, 1993.

\bibitem{papadimitriou1979serial}
C.~H. Papadimitriou.
The serializability of concurrent database updates.
\emph{J.~ACM}, 26(4), 1979.

\bibitem{berenson1995critique}
H.~Berenson, P.~Bernstein, J.~Gray, J.~Melton, E.~O'Neil, and P.~O'Neil.
A critique of ANSI SQL isolation levels.
In \emph{SIGMOD}, 1995.

\bibitem{bernstein1981concurrency}
P.~A. Bernstein and N.~Goodman.
Concurrency control in distributed database systems.
\emph{ACM Comput.\ Surv.}, 13(2), 1981.

\bibitem{denning1976lattice}
D.~E. Denning.
A lattice model of secure information flow.
\emph{Commun.\ ACM}, 19(5), 1976.

\bibitem{myers1997jflow}
A.~C. Myers and B.~Liskov.
A decentralized model for information flow control.
In \emph{SOSP}, 1997.

\bibitem{greshake2023not}
K.~Greshake, S.~Abdelnabi, S.~Mishra, C.~Endres, T.~Holz, and M.~Fritz.
Not what you've signed up for: compromising real-world LLM-integrated
applications with indirect prompt injection.
In \emph{AISec}, 2023.

\bibitem{zhan2024injecagent}
Q.~Zhan, Z.~Liang, Z.~Ying, and D.~Kang.
InjecAgent: benchmarking indirect prompt injections in tool-integrated
large language model agents.
In \emph{Findings of ACL}, 2024.

\bibitem{yao2023react}
S.~Yao, J.~Zhao, D.~Yu, N.~Du, I.~Shafran, K.~Narasimhan, and Y.~Cao.
ReAct: synergizing reasoning and acting in language models.
In \emph{ICLR}, 2023.

\bibitem{agache2020firecracker}
A.~Agache, M.~Brooker, A.~Iordache, A.~Liguori, R.~Neugebauer, P.~Piwonka,
and D.-M. Popa.
Firecracker: lightweight virtualization for serverless applications.
In \emph{NSDI}, 2020.

\bibitem{laurie2014ct}
B.~Laurie.
Certificate transparency.
\emph{Commun.\ ACM}, 57(10), 2014.

\bibitem{lamport2002specifying}
L.~Lamport.
\emph{Specifying Systems: The TLA+ Language and Tools for Hardware and
Software Engineers}.
Addison-Wesley, 2002.

\bibitem{yu1999model}
Y.~Yu, P.~Manolios, and L.~Lamport.
Model checking TLA+ specifications.
In \emph{CHARME}, 1999.

\bibitem{newcombe2015amazon}
C.~Newcombe, T.~Rath, F.~Zhang, B.~Munteanu, M.~Brooker, and M.~Deardeuff.
How Amazon Web Services uses formal methods.
\emph{Commun.\ ACM}, 58(4), 2015.

\bibitem{demillo1978mutation}
R.~A. DeMillo, R.~J. Lipton, and F.~G. Sayward.
Hints on test data selection: help for the practicing programmer.
\emph{IEEE Computer}, 11(4), 1978.

\bibitem{nelsen2006copulas}
R.~B. Nelsen.
\emph{An Introduction to Copulas}, 2nd ed.
Springer, 2006.

\bibitem{boole1854laws}
G.~Boole.
\emph{An Investigation of the Laws of Thought}.
Walton and Maberly, 1854.

\bibitem{hanley1983rule}
J.~A. Hanley and A.~Lippman-Hand.
If nothing goes wrong, is everything all right? Interpreting zero
numerators.
\emph{JAMA}, 249(13), 1983.

\end{thebibliography}

\end{document}